---
title: "Beer Analysis Paper"
author: "Chance Robinson and Thomas Karba"
date: "10/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction

The analysis outlined in this paper seeks to address several questions of interest from the US Craft Beers industry.  There were over 2,410 beers across 558 separate breweries identified in the study.  Although the commissioned study is limited in scope and provides only a small sample of the available attributes, we are confident that it highlights some of the more obvious trends and regional differences in beer tastes.  Two important factors in the flavor of any beer are the percentage of alcohol as well as the bitterness.  We will delve into the specifics of this relationship among beer styles and across different regions of the United States in an attempt to gain better insights into market opportunities.

# Exploratory Data Analysis

## Library Imports


```{r library-imports, quietely=TRUE, warn.conflicts=FALSE}

library(tidyverse)
library(knitr)
library(kableExtra)
library(maps)
library(mapproj)
library(class)
library(caret)
library(e1071)

```


## Load the csv data

```{r load-data}

beers <- read.csv("../../data/Beers.csv", stringsAsFactors=FALSE)
breweries <- read.csv("../../data/Breweries.csv", stringsAsFactors=FALSE)

```

## Default Output

### Beers


```{r head-beers}

# head(beers)
# summary(beers)
# str(beers)

kable(head(beers)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))


```

### Data Dictionary

  Column Name      |Type                  |Description
  -----------------|--------------------- |------------------------------------------------
  1.  Name         |    Character	        | Beer name
  2.  Beer_ID      |    Integer 		      | Unique identifier for beer
  3.  ABV          |    Double 		        | Alcohol by volume
  4.  IBU          |    Integer 		      | International Bitterness Units
  5.  Brewery_id   |    Integer 		      | Foreign key for brewery data set
  6.  Style        |    Character 			  | Beer type
  7.  Ounces       |    Double 			      | Number of ounces per serving



### Breweries

```{r head-breweries}

# head(breweries)
# summary(breweries)
# str(breweries)


kable(head(breweries)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

### Data Dictionary

  Column Name      |Type                  |Description
  -----------------|--------------------- |------------------------------------------------
  1.  Brew_ID      |    Integer	          | Unique identifier for brewery
  2.  Name         |    Integer 		      | Brewery name
  3.  City         |    Character 		    | City of brewery
  4.  State        |    Character 		    | State of brewery


## Clone data sets

* Create a pristine copy of the originals so that further refinements of the columns names, values, etc.. can be performed

### Beers Copy

```{r clone-and-transform-beers}

df_beers <- beers
colnames(df_beers)
names(df_beers) <- c("beer.name", "beer.id", "beer.abv", "beer.ibu", "beer.brewery.id", "beer.style", "beer.ounces")
# head(df_beers)

```

### Breweries Copy

```{r clone-and-transform-breweries}

df_breweries <- breweries
colnames(df_breweries)
names(df_breweries) <- c("brewery.id", "brewery.name", "brewery.city", "brewery.state.abb")

# The state column has an extra white space at the end of the string
df_breweries <- df_breweries %>%
  mutate(brewery.state.abb = trimws(brewery.state.abb))

# head(df_breweries)

```

## State Lookup

* Washington D.C. added as an entry

```{r state-lookup}
#makes a data frame with State abbreviation, name  and region
# df_state_lookup = data.frame(state.abb, state.name, state.region, stringsAsFactors=FALSE) 
df_state_lookup = data.frame(state.abb, state.name, state.region, state.x77, stringsAsFactors=FALSE) 
# df_state_lookup

# add the District of Columbia as a lookup value
df_state_lookup <- rbind(df_state_lookup, 'DC' = c("DC", "Washington D.C.", "South"))
# head(df_state_lookup[1:5])


```



# 1.  Brewery Count by State

Question One:  How many breweries are present in each state?

* Note that all 50 states have at least one brewery, in addition to the District of Columbia (Washington D.C.).

```{r brewery-count}

df_breweries_state <- merge(df_breweries, df_state_lookup[1:5], by.x = "brewery.state.abb", by.y = "state.abb", all.x = TRUE)
# df_breweries_state

df_breweries_count_by_state <- df_breweries_state %>%
  count(state.name, sort = TRUE)

# head(df_breweries_count_by_state)

kable(df_breweries_count_by_state) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))


df_breweries_state %>%
  group_by(brewery.state.abb) %>%
  filter(n() > 25) %>%
  ggplot(aes(x = brewery.state.abb, color=brewery.state.abb)) + 
  geom_bar() +
  geom_text(stat='count', aes(label=..count..), vjust=-1) +
  ggtitle("Brewery Count by State (> 25)") +
  labs(x = "State Abbreviation", y = "Count") +
  theme(plot.title = element_text(hjust = 0.5)) + lims(y=c(0,50)) +
  # guides(fill=guide_legend(title=NULL)) 
  theme(legend.position = "none")


df_breweries_state %>%
  group_by(brewery.state.abb) %>%
  filter(n() < 5) %>%
  ggplot(aes(x = brewery.state.abb, color=brewery.state.abb)) + 
  geom_bar() +
  geom_text(stat='count', aes(label=..count..), vjust=-1) +
  ggtitle("Brewery Count by State (< 5)") +
  labs(x = "State Abbreviation", y = "Count") +
  theme(plot.title = element_text(hjust = 0.5)) + lims(y=c(0,5)) +
  # guides(fill=guide_legend(title=NULL)) 
  theme(legend.position = "none")



```


```{r merge-beer-breweries}

df_merged <- merge(df_beers, df_breweries_state, by.x = "beer.brewery.id", by.y = "brewery.id")
# df_merged

```



# 2.  Merge beer and breweries

Question Two:  Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.  (RMD only, this does not need to be included in the presentation or the deck.)


## First 6 observations of merged data sets

```{r merged-head}

# head(df_merged)

kable(head(df_merged)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

## Last 6 observations of merged data sets

```{r merged-tail}

# tail(df_merged)

kable(tail(df_merged)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))


```


# 3.  Missing Values

Question Three:  Address the missing values in each column

* The ABV and IBU columns have 64 and 1005 NA values respectively
* The Style column has blank or empty values for 5 records

The IBU column in particular, has a relatively high percentage of missing values.  This may complicate our analysis when comparing the data point to other fields that may have many more observations.

```{r merged-missing-values}

df_has_na_values <- any(is.na(df_merged))
df_has_na_values

# which columns have missing values
df_na_columns = which(colSums(is.na(df_merged)) > 0)
df_na_columns

df_na_abv <- df_merged %>%
  filter(is.na(beer.abv))

na_abv_count <- dim(df_na_abv)[1]

df_na_ibu <- df_merged %>%
  filter(is.na(beer.ibu))

na_ibu_count <- dim(df_na_ibu)[1]


allmisscols <- sapply(df_merged, function(x) any(x == '' ))

# allmisscols

df_na_style <- df_merged %>%
  filter(beer.style=='')

na_style_count <- dim(df_na_style)[1]


total_count <- dim(df_merged)[1]


```

  Column Name      |Count                 |% Missing
  -----------------|--------------------- |------------------------------------------------
  1.  ABV          |    `r na_abv_count`	| `r round((na_abv_count / total_count * 100),2)`
  2.  IBU          |    `r na_ibu_count`	| `r round((na_ibu_count / total_count * 100),2)`
  3.  Style        |    `r na_style_count`| `r round((na_style_count / total_count * 100),2)`


# 4.  Barplot of median values

Question Four:  Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.

## Alcohol by Volume

* Kentucky has the highest median ABV of any state in the dataset
* Utah has the lowest median ABV

```{r merged-median-abv-barplot}

# df_merged_abv_clean <- df_merged %>%
#   filter(!is.na(beer.abv)) %>%
#   group_by(state.name) %>%
#   summarize(beers.abv.median = median(beer.abv), count = n()) %>%
#   arrange(desc(beers.abv.median))

# head(df_merged_abv_clean)

# df_merged_abv_clean %>%
#   ggplot(aes(x = reorder(state.name, -beers.abv.median), y=as.numeric(beers.abv.median))) +
#   geom_bar(stat = "identity", position = "dodge") +
#   ggtitle("Bar Plot of median ABV per State") +
#   labs(x = "State", y = "Median") +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1))

df_merged %>%
  select(state.name, beer.abv) %>%
  remove_missing(na.rm = TRUE) %>%
  group_by(state.name) %>%
  mutate(beers.abv.median = median(beer.abv)) %>%
  ggplot(aes(x = reorder(state.name, -beers.abv.median), y=beers.abv.median)) +
  geom_bar(stat = "identity", position = "dodge") +
  ggtitle("Bar Plot of median ABV per State") +
  labs(x = "State", y = "Median") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


## International Bitterness Units

* Maine has the highest median IBU
* Wisconsin has the lowest


```{r merged-median-ibu-barplot}

# df_merged_ibu_clean <- df_merged %>%
#   filter(!is.na(beer.ibu)) %>%
#   group_by(state.name) %>%
#   summarize(beers.ibu.median = median(beer.ibu), count = n()) %>%
#   arrange(desc(beers.ibu.median))

# head(df_merged_ibu_clean)

# df_merged_ibu_clean %>%
#   ggplot(aes(x = reorder(state.name, -beers.ibu.median), y=as.numeric(beers.ibu.median))) +
#   geom_bar(stat = "identity", position = "dodge") +
#   ggtitle("Bar Plot of median IBU per State") +
#   labs(x = "State", y = "Median") +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1))


df_merged %>%
  select(state.name, beer.ibu) %>%
  remove_missing(na.rm = TRUE) %>%
  group_by(state.name) %>%
  mutate(beers.ibu.median = median(beer.ibu)) %>%
  ggplot(aes(x = reorder(state.name, -beers.ibu.median), y=beers.ibu.median)) +
  geom_bar(stat = "identity", position = "dodge") +
  ggtitle("Bar Plot of median IBU per State") +
  labs(x = "State", y = "Median") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```



# 5.  Highest Median Values

Question Five:  Which state has the ...

- maximum alcoholic (ABV) beer?   Colorado
- most bitter (IBU) beer?  Oregon

There are different measures of center, etc.. that might be useful in determining state trends.  As you can see from the tables below that depending on the statistic of interest the answer may vary.



## Alcohol by Volume


```{r merged-median-abv-table}

### Measures of Center
df_merged_abv_max <- df_merged %>%
  filter(!is.na(beer.abv)) %>%
  group_by(state.name) %>%
  summarize(beers.abv.max = max(beer.abv), count = n()) %>%
  arrange(desc(beers.abv.max))
  #top_n(10)

df_merged_abv_median <- df_merged %>%
  filter(!is.na(beer.abv)) %>%
  group_by(state.name) %>%
  summarize(beers.abv.median = median(beer.abv), count = n()) %>%
  arrange(desc(beers.abv.median))
  #top_n(10)

df_merged_abv_mean <- df_merged %>%
  filter(!is.na(beer.abv)) %>%
  group_by(state.name) %>%
  summarize(beers.abv.mean = mean(beer.abv), count = n()) %>%
  arrange(desc(beers.abv.mean))
  # top_n(10)

max_abv <- head(df_merged_abv_max, 1)
median_abv <- head(df_merged_abv_median, 2)
mean_abv <- head(df_merged_abv_mean,1)


head(df_merged_abv_max)
head(df_merged_abv_median)
head(df_merged_abv_mean)


```

  Top ABV Statistics     | State               |Value  
  -----------------------|---------------------|------
  1.  Max                |  `r max_abv[1,1]`	 | `r max_abv[1,2]`     
  2.  Median             |  `r median_abv[1,1]`| `r median_abv[1,2]`     
  2.  Median             |  `r median_abv[2,1]`| `r median_abv[2,2]` 
  3.  Mean               |  `r mean_abv[1]`	   | `r mean_abv[2]`



## International Bitterness Units

```{r merged-median-ibu-table}
  
### Measures of Center
df_merged_ibu_max <- df_merged %>%
  filter(!is.na(beer.ibu)) %>%
  group_by(state.name) %>%
  summarize(beers.ibu.max = max(beer.ibu), count = n()) %>%
  arrange(desc(beers.ibu.max))
  #top_n(10)

df_merged_ibu_median <- df_merged %>%
  filter(!is.na(beer.ibu)) %>%
  group_by(state.name) %>%
  summarize(beers.ibu.median = median(beer.ibu), count = n()) %>%
  arrange(desc(beers.ibu.median))
  #top_n(10)

df_merged_ibu_mean <- df_merged %>%
  filter(!is.na(beer.ibu)) %>%
  group_by(state.name) %>%
  summarize(beers.ibu.mean = mean(beer.ibu), count = n()) %>%
  arrange(desc(beers.ibu.mean))
  #top_n(10)

max_ibu <- head(df_merged_ibu_max, 1)
median_ibu <- head(df_merged_ibu_median, 1)
mean_ibu <- head(df_merged_ibu_mean,1)

head(df_merged_ibu_max)
head(df_merged_ibu_median)
head(df_merged_ibu_mean)

```

  Top IBU Statistics     | State               |Value  
  -----------------------|---------------------|------
  1.  Max                |  `r max_ibu[1,1]`	 | `r max_ibu[1,2]`     
  2.  Median             |  `r median_ibu[1,1]`| `r median_ibu[1,2]`     
  3.  Mean               |  `r mean_ibu[1]`	   | `r mean_ibu[2]`




# 6.  Alcohol by Volume Summary

## ABV Histogram

Question Six:  Comment on the summary statistics and distribution of the ABV variable.

* The distribution of the ABV data is slightly right skewed, with most values falling within ~0.014 units of the mean of ~0.06.


```{r merged-median-abv-histogram}


df_merged_abv_cleaned <- df_merged %>%
  select(state.name, beer.abv) %>%
  remove_missing(na.rm = TRUE)

summary(df_merged_abv_cleaned)

abv_min <- min(df_merged_abv_cleaned$beer.abv)
abv_max <- max(df_merged_abv_cleaned$beer.abv)
abv_median <- median(df_merged_abv_cleaned$beer.abv)
abv_mean <- mean(df_merged_abv_cleaned$beer.abv)
abv_sd <- sd(df_merged_abv_cleaned$beer.abv)


df_merged_abv_cleaned %>%
  ggplot(aes(beer.abv)) +
  geom_histogram() +
  ggtitle("Histogram of median ABV") +
  labs(x = "Median", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


```

## ABV Summary Statistics

  ABV Summary Statistics|Value                 
  -----------------------|--------------------- 
  1.  Minimum            |    `r abv_min`	      
  2.  Maximum            |    `r abv_max`	      
  3.  Median             |    `r abv_median`	 
  4.  Mean               |    `r round(abv_mean,4)`	     
  5.  Standard Deviation |    `r round(abv_sd,4)`	     



# ABV vs IBU

Question Seven:  Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer.

* There appears to be a linear relationship between a beer's International Bitterness Units (IBU) and the Alcohol by volume (ABV).  The stronger a beer is, the more likely it is to rate higher on the bitterness scale.

## Scatterplot

```{r merged-median-abv-ibu-scatterplot}

df_merged %>%
  select(beer.abv, beer.ibu) %>%
  remove_missing(na.rm = TRUE) %>%
  ggplot(aes(x = beer.abv, y = beer.ibu)) + 
  ggtitle("Scatterplot of IBU by ABV") +
  geom_point(alpha = 0.3) + 
  labs(x = "ABV", y = "IBU") +
  # scale_x_continuous(breaks = seq(from = 0, to = 45, by = 2)) + 
  geom_smooth(method = 'lm')


df_abv_ibu_cor <- df_merged %>%
  select(beer.abv, beer.ibu)


```

## Pearson's Correlation

* Roughly 67% of the variation of a beer's IBU rating can be explained by the ABV for observations which had both attributes to review.

```{r merged-median-abv-ibu-correlation}

r <- cor(as.matrix(df_abv_ibu_cor), use="complete.obs")
r

```


## Correlation Test

* We estimate that this ratio falls within the interval of (0.6408, 0.6984) at the 95% confidence level.

```{r merged-median-abv-ibu-correlation-test}

t <- cor.test(df_abv_ibu_cor$beer.abv, df_abv_ibu_cor$beer.ibu, use="complete.obs")
t
# t$estimate

```


# 8.  KNN

Question Eight:  Budweiser would also like to investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale (any beer with “Ale” in its name other than IPA).  You decide to use KNN clustering to investigate this relationship.  Provide statistical evidence one way or the other. You can of course assume your audience is comfortable with percentages … KNN is very easy to understand.  

 

```{r add-ipa-nonipa}

df_merged$beer.ipa.vs.ale <- ''

df_merged$Population <- as.numeric(df_merged$Population)
df_merged$Income <- as.numeric(df_merged$Income)

# head(df_merged)

```


* Group by Beer Style and get counts for each

```{r all-styles}

df_merged %>%
  count(beer.style, sort = TRUE) %>%
  top_n(10)

# kable(df_merged %>%
#   count(beer.style, sort = TRUE)) %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

* Five beers have not style available, one of which appears to be an Ale.

```{r no-style}

df_merged %>%
  filter(beer.style == '')

```

### IPAs

* We were able to clasify 560 beers as being IPAs

- American Double / Imperial IPA
- American IPA
- Belgian IPA
- English India Pale Ale (IPA)

```{r ipa-beers}

df_merged %>%
  filter(beer.style %in% c("American Double / Imperial IPA", "American IPA", "Belgian IPA", "English India Pale Ale (IPA)")) %>%
  count(beer.style)

df_ipa <- df_merged %>%
  filter(beer.style %in% c("American Double / Imperial IPA", "American IPA", "Belgian IPA", "English India Pale Ale (IPA)")) %>%
  mutate(beer.ipa.vs.ale = 'IPA')

# head(df_ipa)
```

### Other Ales

* We were able to clasify 964 beers as being Ales

```{r other-ales}

# Kilt Lifter Scottish-Style Ale (1635)
df_merged[946,]

df_merged %>%
  filter(!beer.style %in% c("American Double / Imperial IPA", "American IPA", "Belgian IPA", "English India Pale Ale (IPA)")) %>%
  filter(str_detect(beer.style, "Ale") | beer.id==1635) %>%
  count(beer.style)

df_ale <- df_merged %>%
  filter(!beer.style %in% c("American Double / Imperial IPA", "American IPA", "Belgian IPA", "English India Pale Ale (IPA)")) %>%
  filter(str_detect(beer.style, "Ale") | beer.id==1635) %>%
  mutate(beer.ipa.vs.ale = 'Ale')

# head(df_ale)

```

### Rowbind the ipa and ale data frames

```{r bind-ipa-ale}

total_beers_count <- dim(df_merged)[1]
ipa_count <- dim(df_ipa)[1]
ale_count <- dim(df_ale)[1]

df_combined_ale <- rbind(df_ipa, df_ale)

ipa_ale_count <- dim(df_combined_ale)[1]


# head(df_combined_ale)

```

  Column Name      |Count                 |% of Total
  -----------------|--------------------- |------------------------------------------------
  1.  IPA          |    `r ipa_count`	    | `r round((ipa_count / total_beers_count * 100),2)`
  2.  Other Ales   |    `r ale_count`	    | `r round((ale_count / total_beers_count * 100),2)`
  3.  Ales + IPAs  |    `r ipa_ale_count` | `r round((ipa_ale_count / total_beers_count * 100),2)`
  

### ABV Boxplot

```{r ipa-vs-other-ale-abv-boxplot}

df_combined_ale %>% 
  filter(!is.na(beer.abv)) %>%
  ggplot(aes(x = beer.ipa.vs.ale, y = beer.abv, fill = beer.ipa.vs.ale)) + 
  geom_boxplot() + 
  ggtitle("Boxplot of Alcohol by Volume (IPAs vs. Other Ales)") + 
  xlab("Style") + 
  ylab("ABV")

```


### IBU Boxplot

```{r ipa-vs-other-ale-ibu-boxplot}

df_combined_ale %>% 
  filter(!is.na(beer.ibu)) %>%
  ggplot(aes(x = beer.ipa.vs.ale, y = beer.ibu, fill = beer.ipa.vs.ale)) + 
  geom_boxplot() + 
  ggtitle("Boxplot of International Bitterness Units (IPAs vs. Other Ales)") + 
  xlab("Style") + 
  ylab("IBU")

```


## Scatterplot

```{r classify-dataframe}
df_classify <- df_combined_ale %>% 
  filter(!is.na(beer.abv) & !is.na(beer.ibu))

df_classify %>%
  group_by(beer.ipa.vs.ale) %>%
  summarize(beers.abv.median = median(beer.abv), beers.ibu.median = median(beer.ibu), count = n())


df_classify %>% 
  ggplot(aes(x = beer.abv, y = beer.ibu, color =  beer.ipa.vs.ale)) + 
  xlab("ABV") + ylab("IBU") + 
  geom_point()


# df = data.frame(beer.abv = .07, beer.ibu = 50 )
# knn(df_combined_ale_cleaned[,c(4,5)], df, df_combined_ale_cleaned$beer.ipa.vs.ale, k = 5, prob = TRUE)


```


```{r seed-data}

set.seed(1234)
split.perc = .70

# summary(df_classify)


# standardize values
df_classify$z.abv = scale(df_classify$beer.abv)
df_classify$z.ibu = scale(df_classify$beer.ibu)


train.indices = sample(1:dim(df_classify)[1],round(split.perc * dim(df_classify)[1]))

train = df_classify[train.indices,]
test = df_classify[-train.indices,]

dim(df_classify) # 939
dim(train) # 657
dim(test) #282



# original scale
classifications = knn(train[,c(4,5)],test[,c(4,5)], train$beer.ipa.vs.ale, prob = TRUE, k = 6)
# confusionMatrix(table(test$beer.ipa.vs.ale,classifications))

# standardized
classifications = knn(train[,c(16,17)],test[,c(16,17)], train$beer.ipa.vs.ale, prob = TRUE, k = 16)
confusionMatrix(table(test$beer.ipa.vs.ale,classifications))


```


### Loop for many k and one training / test partition

```{r loop-many-k}
acc = data.frame(accuracy = numeric(30), k = numeric(30))

for(i in 1:30)
{
  classifications = knn(train[,c(16,17)],test[,c(16,17)], train$beer.ipa.vs.ale, prob = TRUE, k = i)
  table(test$beer.ipa.vs.ale, classifications)
  CM = confusionMatrix(table(test$beer.ipa.vs.ale,classifications))
  acc$accuracy[i] = CM$overall[1]
  acc$k[i] = i
}

plot(acc$k,acc$accuracy, type = "l", xlab = "k", ylab="accuracy")

```

## Loop for many k and many training / test partitions

```{r loop-many-k-many-train-test}

iterations = 100
numks = 30


masterAcc = matrix(nrow = iterations, ncol = numks)

for(j in 1:iterations)
{
  train.indices = sample(1:dim(df_classify)[1],round(split.perc * dim(df_classify)[1]))
  train = df_classify[train.indices,]
  test = df_classify[-train.indices,]
  for(i in 1:numks)
  {
    classifications = knn(train[,c(16,17)],test[,c(16,17)], train$beer.ipa.vs.ale, prob = TRUE, k = i)
    table(test$beer.ipa.vs.ale, classifications)
    CM = confusionMatrix(table(test$beer.ipa.vs.ale,classifications))
    masterAcc[j,i] = CM$overall[1]
  }
  
}

MeanAcc = colMeans(masterAcc)

plot(seq(1,numks,1),MeanAcc, type = "l")

which.max(MeanAcc)
max(MeanAcc)



```



# Headmap of US Craft Beers by State


Question Nine:  Knock their socks off!  Find one other useful inference from the data that you feel Budweiser may be able to find value in.  You must convince them why it is important and back up your conviction with appropriate statistical evidence. 


* More craft beers are made in the West than any other region.  Colorado, California and Oregon for example are all in the top 5 as far as having the most craft beers made in-state.


```{r merged-count-beers-by-state-heatmap}


#count up the occurance of each state. 
df_beer_count_by_state <- df_merged %>%
  count(state.name, sort = TRUE)

head(df_beer_count_by_state)


colnames(df_beer_count_by_state)[1] = "State" #change "state.name" to "State"
colnames(df_beer_count_by_state)[2] = "Count" #change "n" to "Count"

# df_abv_map_data1

df_beer_count_by_state$region <- tolower(df_beer_count_by_state$State)
df_beer_count_by_state2 <- df_beer_count_by_state[-1]
states <- map_data("state")
map.df <- merge(states, df_beer_count_by_state2, by="region", all.x=T)
map.df <- map.df[order(map.df$order),]
ggplot(map.df, aes(x=long,y=lat,group=group))+
  geom_polygon(aes(fill=Count))+
  geom_path()+ 
  scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey90")+ggtitle("US Craft Beer Count by State (Heatmap)")+
coord_map()
```

```{r merged-count-beers-by-region-barplot}


#count up the occurance of each state. 
df_beer_count_by_region <- df_merged %>%
  count(state.region, sort = TRUE) %>%
  ggplot(aes(x = reorder(state.region, -n), y=n)) +
  geom_bar(stat = "identity", position = "dodge") +
  ggtitle("Bar Plot of US Craft Beers by Region") +
  labs(x = "State", y = "Median") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

df_beer_count_by_region

# df_merged %>%
#   count(state.region, sort = TRUE) %>%
#   ggplot(aes(x = reorder(state.region, -n), y=n)) +
#   geom_bar(stat = "identity", position = "dodge") +
#   # ggtitle("Brewery Count by Region") +
#   labs(x = "Region", y = "Count") +
#   ggtitle("US Craft Beer Count by Region")


```


# Conclusion




The analysis outlined in this paper seeks to address several questions of interest from the US Craft Beers industry.  There were over 2,410 beers across 558 separate breweries identified in the study.  Although the commissioned study is limited in scope and provides only a small sample of the available attributes, we are confident that it highlights some of the more obvious trends and regional differences in beer tastes.  Two important factors in the flavor of any beer are the percentage of alcohol as well as the bitterness.  We will delve into the specifics of this relationship among beer styles and across different regions of the United States in an attempt to gain better insights into market opportunities.

