---
title: "Beer Analysis Paper"
author: "Chance Robinson and Thomas Karba"
date: "10/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exploratory Data Analysis

## Library Imports


```{r library-imports, quietely=TRUE, warn.conflicts=FALSE}

library(tidyverse)
library(knitr)
library(kableExtra)
library(maps)
library(mapproj)

```


## Load the csv data

```{r load-data}

beers <- read.csv("../../data/Beers.csv", stringsAsFactors=FALSE)
breweries <- read.csv("../../data/Breweries.csv", stringsAsFactors=FALSE)

```

## Default Output

### Beers


```{r head-beers}

head(beers)
# summary(beers)
# str(beers)


```

#### Data Dictionary

  Column Name      |Type                  |Description
  -----------------|--------------------- |------------------------------------------------
  1.  Name         |    Character	        | Beer name
  2.  Beer_ID      |    Integer 		      | Unique identifier for beer
  3.  ABV          |    Double 		        | Alcohol by volume
  4.  IBU          |    Integer 		      | International Bitterness Units
  5.  Brewery_id   |    Integer 		      | Foreign key for brewery data set
  6.  Style        |    Character 			  | Beer type
  7.  Ounces       |    Double 			      | Number of ounces per serving



### Breweries

```{r head-breweries}

head(breweries)
# summary(breweries)
# str(breweries)

```

#### Data Dictionary

  Column Name      |Type                  |Description
  -----------------|--------------------- |------------------------------------------------
  1.  Brew_ID      |    Integer	          | Unique identifier for brewery
  2.  Name         |    Integer 		      | Brewery name
  3.  City         |    Character 		    | City of brewery
  4.  State        |    Character 		    | State of brewery


## Clone data sets

* Create a pristine copy of the originals so that further refinements of the columns names, values, etc.. can be performed

### Beers

```{r clone-and-transform-beers}

df_beers <- beers
colnames(df_beers)
names(df_beers) <- c("beer.name", "beer.id", "beer.abv", "beer.ibu", "beer.brewery.id", "beer.style", "beer.ounces")
head(df_beers)

```

### Breweries

```{r clone-and-transform-breweries}

df_breweries <- breweries
colnames(df_breweries)
names(df_breweries) <- c("brewery.id", "brewery.name", "brewery.city", "brewery.state.abb")

# The state column has an extra white space at the end of the string
df_breweries <- df_breweries %>%
  mutate(brewery.state.abb = trimws(brewery.state.abb))

head(df_breweries)

```

## State Lookup

- Washington D.C. added as an entry

```{r state-lookup}
#makes a data frame with State abbreviation, name  and region
df_state_lookup = data.frame(state.abb, state.name, state.region, stringsAsFactors=FALSE) 
# df_state_lookup

# add the District of Columbia as a lookup value
df_state_lookup <- rbind(df_state_lookup, 'DC' = c("DC", "Washington D.C.", "South"))
head(df_state_lookup)


```



## Brewery Count by State

1.  How many breweries are present in each state?

* Note that all 50 states have counts and that the District of Columbia is also included to make a total of 51

```{r brewery-count}

df_breweries_state <- merge(df_breweries, df_state_lookup, by.x = "brewery.state.abb", by.y = "state.abb", all.x = TRUE)
# df_breweries_state

df_breweries_count_by_state <- df_breweries_state %>%
  count(state.name, sort = TRUE)

# head(df_breweries_count_by_state)

kable(df_breweries_count_by_state) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))


df_breweries_state %>%
  group_by(brewery.state.abb) %>%
  filter(n() > 25) %>%
  ggplot(aes(x = brewery.state.abb, color=brewery.state.abb)) + 
  geom_bar() +
  geom_text(stat='count', aes(label=..count..), vjust=-1) +
  ggtitle("Brewery Count by State (> 25)") +
  labs(x = "State Abbreviation", y = "Count") +
  theme(plot.title = element_text(hjust = 0.5)) + lims(y=c(0,50)) +
  # guides(fill=guide_legend(title=NULL)) 
  theme(legend.position = "none")


df_breweries_state %>%
  group_by(brewery.state.abb) %>%
  filter(n() < 5) %>%
  ggplot(aes(x = brewery.state.abb, color=brewery.state.abb)) + 
  geom_bar() +
  geom_text(stat='count', aes(label=..count..), vjust=-1) +
  ggtitle("Brewery Count by State (< 5)") +
  labs(x = "State Abbreviation", y = "Count") +
  theme(plot.title = element_text(hjust = 0.5)) + lims(y=c(0,5)) +
  # guides(fill=guide_legend(title=NULL)) 
  theme(legend.position = "none")



```


```{r merge-beer-breweries}

df_merged <- merge(df_beers, df_breweries_state, by.x = "beer.brewery.id", by.y = "brewery.id")
# df_merged

```



## Merge beer and breweries

2.  Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.  (RMD only, this does not need to be included in the presentation or the deck.)


### First 6 observations of merged data sets
```{r merged-head}

head(df_merged)

```

### Last 6 observations of merged data sets

```{r merged-tail}

tail(df_merged)


```


## Missing Values

3.   Address the missing values in each column.

* The ABV and IBU columns have NA values
* The Style column as blank values

- 62 for ABV
- 1005 for IBU (this is a very high percentage of the overall rows)

```{r merged-missing-values}

df_has_na_values <- any(is.na(df_merged))
df_has_na_values

# which columns have missing values
df_na_columns = which(colSums(is.na(df_merged)) > 0)
df_na_columns

df_na_abv <- df_merged %>%
  filter(is.na(beer.abv))

na_abv_count <- dim(df_na_abv)[1]

df_na_ibu <- df_merged %>%
  filter(is.na(beer.ibu))

na_ibu_count <- dim(df_na_ibu)[1]


allmisscols <- sapply(df_merged, function(x) any(x == '' ))

# allmisscols

df_na_style <- df_merged %>%
  filter(beer.style=='')

na_style_count <- dim(df_na_style)[1]


total_count <- dim(df_merged)[1]


```

  Column Name      |Count                 |% Missing
  -----------------|--------------------- |------------------------------------------------
  1.  ABV          |    `r na_abv_count`	| `r na_abv_count / total_count * 100`
  2.  IBU          |    `r na_ibu_count`	| `r na_ibu_count / total_count * 100`
  3.  Style        |    `r na_style_count`| `r na_style_count / total_count * 100`


## Barplot of median values

4.   Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.

### ABV

```{r merged-median-abv-barplot}

# df_merged_abv_clean <- df_merged %>%
#   filter(!is.na(beer.abv)) %>%
#   group_by(state.name) %>%
#   summarize(beers.abv.median = median(beer.abv), count = n()) %>%
#   arrange(desc(beers.abv.median))

# head(df_merged_abv_clean)

# df_merged_abv_clean %>%
#   ggplot(aes(x = reorder(state.name, -beers.abv.median), y=as.numeric(beers.abv.median))) +
#   geom_bar(stat = "identity", position = "dodge") +
#   ggtitle("Bar Plot of median ABV per State") +
#   labs(x = "State", y = "Median") +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1))

df_merged %>%
  select(state.name, beer.abv) %>%
  remove_missing(na.rm = TRUE) %>%
  group_by(state.name) %>%
  mutate(beers.abv.median = median(beer.abv)) %>%
  ggplot(aes(x = reorder(state.name, -beers.abv.median), y=beers.abv.median)) +
  geom_bar(stat = "identity", position = "dodge") +
  ggtitle("Bar Plot of median ABV per State") +
  labs(x = "State", y = "Median") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


### IBU

```{r merged-median-ibu-barplot}

# df_merged_ibu_clean <- df_merged %>%
#   filter(!is.na(beer.ibu)) %>%
#   group_by(state.name) %>%
#   summarize(beers.ibu.median = median(beer.ibu), count = n()) %>%
#   arrange(desc(beers.ibu.median))

# head(df_merged_ibu_clean)

# df_merged_ibu_clean %>%
#   ggplot(aes(x = reorder(state.name, -beers.ibu.median), y=as.numeric(beers.ibu.median))) +
#   geom_bar(stat = "identity", position = "dodge") +
#   ggtitle("Bar Plot of median IBU per State") +
#   labs(x = "State", y = "Median") +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1))


df_merged %>%
  select(state.name, beer.ibu) %>%
  remove_missing(na.rm = TRUE) %>%
  group_by(state.name) %>%
  mutate(beers.ibu.median = median(beer.ibu)) %>%
  ggplot(aes(x = reorder(state.name, -beers.ibu.median), y=beers.ibu.median)) +
  geom_bar(stat = "identity", position = "dodge") +
  ggtitle("Bar Plot of median IBU per State") +
  labs(x = "State", y = "Median") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```



## Highest Median Values

5.  Which state has the highest...

- maximum alcoholic (ABV) beer?   Colorado
- most bitter (IBU) beer?  Oregon

### ABV


```{r merged-median-abv-table}

### Measures of Center
df_merged_abv_max <- df_merged %>%
  filter(!is.na(beer.abv)) %>%
  group_by(state.name) %>%
  summarize(beers.abv.max = max(beer.abv), count = n()) %>%
  arrange(desc(beers.abv.max))
  #top_n(10)

df_merged_abv_median <- df_merged %>%
  filter(!is.na(beer.abv)) %>%
  group_by(state.name) %>%
  summarize(beers.abv.median = median(beer.abv), count = n()) %>%
  arrange(desc(beers.abv.median))
  #top_n(10)

df_merged_abv_mean <- df_merged %>%
  filter(!is.na(beer.abv)) %>%
  group_by(state.name) %>%
  summarize(beers.abv.mean = mean(beer.abv), count = n()) %>%
  arrange(desc(beers.abv.mean))
  # top_n(10)

max_abv <- head(df_merged_abv_max, 1)
median_abv <- head(df_merged_abv_median, 2)
mean_abv <- head(df_merged_abv_mean,1)


head(df_merged_abv_max)
head(df_merged_abv_median)
head(df_merged_abv_mean)


```

  Top ABV Statistics     | State               |Value  
  -----------------------|---------------------|------
  1.  Max                |  `r max_abv[1,1]`	 | `r max_abv[1,2]`     
  2.  Median             |  `r median_abv[1,1]`| `r median_abv[1,2]`     
  2.  Median             |  `r median_abv[2,1]`| `r median_abv[2,2]` 
  3.  Mean               |  `r mean_abv[1]`	   | `r mean_abv[2]`



### IBU

```{r merged-median-ibu-table}
  
### Measures of Center
df_merged_ibu_max <- df_merged %>%
  filter(!is.na(beer.ibu)) %>%
  group_by(state.name) %>%
  summarize(beers.ibu.max = max(beer.ibu), count = n()) %>%
  arrange(desc(beers.ibu.max))
  #top_n(10)

df_merged_ibu_median <- df_merged %>%
  filter(!is.na(beer.ibu)) %>%
  group_by(state.name) %>%
  summarize(beers.ibu.median = median(beer.ibu), count = n()) %>%
  arrange(desc(beers.ibu.median))
  #top_n(10)

df_merged_ibu_mean <- df_merged %>%
  filter(!is.na(beer.ibu)) %>%
  group_by(state.name) %>%
  summarize(beers.ibu.mean = mean(beer.ibu), count = n()) %>%
  arrange(desc(beers.ibu.mean))
  #top_n(10)

max_ibu <- head(df_merged_ibu_max, 1)
median_ibu <- head(df_merged_ibu_median, 1)
mean_ibu <- head(df_merged_ibu_mean,1)

head(df_merged_ibu_max)
head(df_merged_ibu_median)
head(df_merged_ibu_mean)

```

  Top IBU Statistics     | State               |Value  
  -----------------------|---------------------|------
  1.  Max                |  `r max_ibu[1,1]`	 | `r max_ibu[1,2]`     
  2.  Median             |  `r median_ibu[1,1]`| `r median_ibu[1,2]`     
  3.  Mean               |  `r mean_ibu[1]`	   | `r mean_ibu[2]`




## ABV Summaary

### Histogram

6.  Comment on the summary statistics and distribution of the ABV variable.

* The histogram is right-skewed with the majority of the beers falling between 0.05 and 0.067 % alcholol by volume.

```{r merged-median-abv-histogram}


df_merged_abv_cleaned <- df_merged %>%
  select(state.name, beer.abv) %>%
  remove_missing(na.rm = TRUE)

summary(df_merged_abv_cleaned)

abv_min <- min(df_merged_abv_cleaned$beer.abv)
abv_max <- max(df_merged_abv_cleaned$beer.abv)
abv_median <- median(df_merged_abv_cleaned$beer.abv)
abv_mean <- mean(df_merged_abv_cleaned$beer.abv)
abv_sd <- sd(df_merged_abv_cleaned$beer.abv)


df_merged_abv_cleaned %>%
  ggplot(aes(beer.abv)) +
  geom_histogram() +
  ggtitle("Histogram of median ABV") +
  labs(x = "Median", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


```

### Summary Statistics

  ABV Summary Statistics|Value                 
  -----------------------|--------------------- 
  1.  Minimum            |    `r abv_min`	      
  2.  Maximum            |    `r abv_max`	      
  3.  Median             |    `r abv_median`	 
  4.  Mean               |    `r abv_mean`	     
  5.  Standard Deviation |    `r abv_sd`	     


```{r merged-median-abv-summary}

abv_min 
abv_max 
abv_median 
abv_mean 
abv_sd 

```


## ABV vs IBU

7.  Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer.

- There appears to be a linear relationship between a beer's International Bitterness Units (IBU) and the Alcohol by volume (ABV).  The stronger a beer is, the more likely it is to rate higher on the bitterness scale.

### Scatterplot

```{r merged-median-abv-ibu-scatterplot}

df_merged %>%
  select(beer.abv, beer.ibu) %>%
  remove_missing(na.rm = TRUE) %>%
  ggplot(aes(x = beer.abv, y = beer.ibu)) + 
  ggtitle("Scatterplot of IBU by ABV") +
  geom_point(alpha = 0.3) + 
  labs(x = "ABV", y = "IBU") +
  # scale_x_continuous(breaks = seq(from = 0, to = 45, by = 2)) + 
  geom_smooth(method = 'lm')


df_abv_ibu_cor <- df_merged %>%
  select(beer.abv, beer.ibu)


```

### Pearson's Correlation
```{r merged-median-abv-ibu-correlation}

r <- cor(as.matrix(df_abv_ibu_cor), use="complete.obs")
r

```


### Correlation Test
```{r merged-median-abv-ibu-correlation-test}

t <- cor.test(df_abv_ibu_cor$beer.abv, df_abv_ibu_cor$beer.ibu, use="complete.obs")
t
# t$estimate

```


## KNN

7.  Budweiser would also like to investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale (any beer with “Ale” in its name other than IPA).  You decide to use KNN clustering to investigate this relationship.  Provide statistical evidence one way or the other. You can of course assume your audience is comfortable with percentages … KNN is very easy to understand.  

- ...



## Headmap of US Craft Beers by State


9. Knock their socks off!  Find one other useful inference from the data that you feel Budweiser may be able to find value in.  You must convince them why it is important and back up your conviction with appropriate statistical evidence. 


- More craft beers are made in the West than any other region.  Colorado, California and Oregon for example are all in the top 5 as far as having the most craft beers made in-state.


```{r merged-count-beers-by-state-heatmap}


#count up the occurance of each state. 
df_beer_count_by_state <- df_merged %>%
  count(state.name, sort = TRUE)

df_beer_count_by_state


colnames(df_beer_count_by_state)[1] = "State" #change "state.name" to "State"
colnames(df_beer_count_by_state)[2] = "Count" #change "n" to "Count"

# df_abv_map_data1

df_beer_count_by_state$region <- tolower(df_beer_count_by_state$State)
df_beer_count_by_state2 <- df_beer_count_by_state[-1]
states <- map_data("state")
map.df <- merge(states, df_beer_count_by_state2, by="region", all.x=T)
map.df <- map.df[order(map.df$order),]
ggplot(map.df, aes(x=long,y=lat,group=group))+
  geom_polygon(aes(fill=Count))+
  geom_path()+ 
  scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey90")+ggtitle("US Craft Beer Count by State (Heatmap)")+
coord_map()
```

```{r merged-count-beers-by-region-barplot}


#count up the occurance of each state. 
df_beer_count_by_region <- df_merged %>%
  count(state.region, sort = TRUE) %>%
  ggplot(aes(x = reorder(state.region, -n), y=n)) +
  geom_bar(stat = "identity", position = "dodge") +
  ggtitle("Bar Plot of US Craft Beers by Region") +
  labs(x = "State", y = "Median") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

df_beer_count_by_region

# df_merged %>%
#   count(state.region, sort = TRUE) %>%
#   ggplot(aes(x = reorder(state.region, -n), y=n)) +
#   geom_bar(stat = "identity", position = "dodge") +
#   # ggtitle("Brewery Count by Region") +
#   labs(x = "Region", y = "Count") +
#   ggtitle("US Craft Beer Count by Region")


```
