---
title: "IPA Beer Comparisons to Other Ale Beers"
author: "Chance Robinson"
date: "10/21/2019"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exploratory Data Analysis

## Library Imports


```{r library-imports, quietely=TRUE, warn.conflicts=FALSE}

library(tidyverse)
library(knitr)
library(kableExtra)

```


## Load the csv data

```{r load-data}

beers <- read.csv("../../../data/Beers.csv", stringsAsFactors=FALSE)
breweries <- read.csv("../../../data/Breweries.csv", stringsAsFactors=FALSE)

```

## Default Output

### Beers


```{r head-beers}

head(beers)
# summary(beers)
# str(beers)

```


### Breweries

```{r head-breweries}

head(breweries)
# summary(breweries)
# str(breweries)

```


## Clone data sets

* Create a pristine copy of the originals so that further refinements of the columns names, values, etc.. can be performed

### Beers

```{r clone-and-transform-beers}

df_beers <- beers
colnames(df_beers)
names(df_beers) <- c("beer.name", "beer.id", "beer.abv", "beer.ibu", "beer.brewery.id", "beer.style", "beer.ounces")
head(df_beers)

```

### Breweries

```{r clone-and-transform-breweries}

df_breweries <- breweries
colnames(df_breweries)
names(df_breweries) <- c("brewery.id", "brewery.name", "brewery.city", "brewery.state.abb")

# The state column has an extra white space at the end of the string
df_breweries <- df_breweries %>%
  mutate(brewery.state.abb = trimws(brewery.state.abb))

head(df_breweries)

```

## State Lookup

- Washington D.C. added as an entry

```{r state-lookup}
#makes a data frame with State abbreviation, name  and region
df_state_lookup = data.frame(state.abb, state.name, state.region, state.x77, stringsAsFactors=FALSE) 
# df_state_lookup

# add the District of Columbia as a lookup value
df_state_lookup <- rbind(df_state_lookup, 'DC' = c("DC", "Washington D.C.", "South"))

head(df_state_lookup[1:5])


```



## Brewery Count by State

1.  How many breweries are present in each state?

* Note that all 50 states have counts and that the District of Columbia is also included to make a total of 51

```{r brewery-count}

df_breweries_state <- merge(df_breweries, df_state_lookup[1:5], by.x = "brewery.state.abb", by.y = "state.abb", all.x = TRUE)

head(df_breweries_state)


```


```{r merge-beer-breweries}

df_merged <- merge(df_beers, df_breweries_state, by.x = "beer.brewery.id", by.y = "brewery.id")
head(df_merged)

```

```{r add-ipa-nonipa}

df_merged$beer.ipa.vs.ale <- ''

head(df_merged)

```


```{r all-styles}

df_merged %>%
  count(beer.style)
```


```{r no-style}

df_merged %>%
  filter(beer.style == '')

```


```{r ipa-beers}

df_merged %>%
  filter(beer.style %in% c("American Double / Imperial IPA", "American IPA", "Belgian IPA", "English India Pale Ale (IPA)")) %>%
  count(beer.style)

df_ipa <- df_merged %>%
  filter(beer.style %in% c("American Double / Imperial IPA", "American IPA", "Belgian IPA", "English India Pale Ale (IPA)")) %>%
  mutate(beer.ipa.vs.ale = 'IPA')


head(df_ipa)
```



```{r other-ales}

# Kilt Lifter Scottish-Style Ale (1635)
df_merged[946,]

df_merged %>%
  filter(!beer.style %in% c("American Double / Imperial IPA", "American IPA", "Belgian IPA", "English India Pale Ale (IPA)")) %>%
  filter(str_detect(beer.style, "Ale") | beer.id==1635) %>%
  count(beer.style)

df_ale <- df_merged %>%
  filter(!beer.style %in% c("American Double / Imperial IPA", "American IPA", "Belgian IPA", "English India Pale Ale (IPA)")) %>%
  filter(str_detect(beer.style, "Ale") | beer.id==1635) %>%
  mutate(beer.ipa.vs.ale = 'Ale')

head(df_ale)

```

```{r head-merged}
# df_merged <- as.factor(df_merged$beer.ipa.vs.ale)
# 
head(df_merged)
```



```{r boundry-words}

head(df_merged) %>%
  mutate(beer.style.words = as.character(str_split(beer.style, boundary("word")))) %>%
  select(beer.style.words)


# str_split(df_merged$beer.style, boundary("word"))

```

```{r regex}

# Ale

# India Pale Ale
IPA <- c("American Double / Imperial IPA", "American IPA", "Belgian IPA", "English India Pale Ale (IPA)")


```



```{r american-pale-ale-abv}

df_merged %>%
  filter(beer.style %in% c("American Pale Ale (APA)")) %>%
  filter(!is.na(beer.abv)) %>%
  summarize(beers.abv.mean = mean(beer.abv), 
            beers.abv.median = median(beer.abv), 
            beers.abv.sd = sd(beer.abv), 
            beers.abv.min = min(beer.abv), 
            beers.abv.max = max(beer.abv), 
            count = n()) 

```


```{r ipa-ibu}

df_merged %>%
  filter(beer.style  %in% c("American Double / Imperial IPA", "American IPA", "Belgian IPA", "English India Pale Ale (IPA)")) %>%
  filter(!is.na(beer.abv)) %>%
  summarize(beers.abv.mean = mean(beer.abv), 
            beers.abv.median = median(beer.abv), 
            beers.abv.sd = sd(beer.abv), 
            beers.abv.min = min(beer.abv), 
            beers.abv.max = max(beer.abv), 
            count = n()) 

```




```{r american-pale-ale-ibu}

df_merged %>%
  filter(beer.style %in% c("American Pale Ale (APA)")) %>%
  filter(!is.na(beer.ibu)) %>%
  summarize(beers.ibu.mean = mean(beer.ibu), 
            beers.ibu.median = median(beer.ibu), 
            beers.ibu.sd = sd(beer.ibu), 
            beers.ibu.min = min(beer.ibu), 
            beers.ibu.max = max(beer.ibu), 
            count = n()) 
```


```{r ale-ibu}

df_merged %>%
  filter(!beer.style %in% c("American Double / Imperial IPA", "American IPA", "Belgian IPA", "English India Pale Ale (IPA)")) %>%
  filter(str_detect(beer.style, "Ale")) %>%
  filter(!is.na(beer.ibu)) %>%
  summarize(beers.ibu.mean = mean(beer.ibu), 
            beers.ibu.median = median(beer.ibu), 
            beers.ibu.sd = sd(beer.ibu), 
            beers.ibu.min = min(beer.ibu), 
            beers.ibu.max = max(beer.ibu), 
            count = n()) 
  

```


```{r bind-ipa-ale}

total_beers_count <- dim(df_merged)[1]
ipa_count <- dim(df_ipa)[1]
ale_count <- dim(df_ale)[1]

df_combined_ale <- rbind(df_ipa, df_ale)

ipa_ale_count <- dim(df_combined_ale)[1]


head(df_combined_ale)

```

  Column Name      |Count                 |% of Total
  -----------------|--------------------- |------------------------------------------------
  1.  IPA          |    `r ipa_count`	    | `r ipa_count / total_beers_count * 100`
  2.  Other Ales   |    `r ale_count`	    | `r ale_count / total_beers_count * 100`
  3.  Ales + IPAs  |    `r ipa_ale_count` | `r ipa_ale_count / total_beers_count * 100`
  
  
```{r ipa-vs-other-ale-abv-boxplot}

df_combined_ale %>% 
  filter(!is.na(beer.abv)) %>%
  ggplot(aes(x = beer.ipa.vs.ale, y = beer.abv, fill = beer.ipa.vs.ale)) + 
  geom_boxplot() + 
  ggtitle("Boxplot of Alcohol by Volume (IPAs vs. Other Ales)") + 
  xlab("Style") + 
  ylab("ABV")

```
  
 
 
```{r ipa-vs-other-ale-ibu-boxplot}

df_combined_ale %>% 
  filter(!is.na(beer.ibu)) %>%
  ggplot(aes(x = beer.ipa.vs.ale, y = beer.ibu, fill = beer.ipa.vs.ale)) + 
  geom_boxplot() + 
  ggtitle("Boxplot of Int. Bitterness Units (IPAs vs. Other Ales)") + 
  xlab("Style") + 
  ylab("IBU")

```
 
```{r}
library(class)
library(caret)
library(e1071)
```


```{r}
df_classify <- df_combined_ale %>% 
  filter(!is.na(beer.abv) & !is.na(beer.ibu))

df_classify %>%
  group_by(beer.ipa.vs.ale) %>%
  summarize(beers.abv.median = median(beer.abv), beers.ibu.median = median(beer.ibu), count = n())


df_classify %>% 
  ggplot(aes(x = beer.abv, y = beer.ibu, color =  beer.ipa.vs.ale)) + 
  xlab("ABV") + ylab("IBU") + 
  geom_point()


# df = data.frame(beer.abv = .07, beer.ibu = 50 )
# knn(df_combined_ale_cleaned[,c(4,5)], df, df_combined_ale_cleaned$beer.ipa.vs.ale, k = 5, prob = TRUE)


```


```{r}

set.seed(1234)
split.perc = .70

# summary(df_classify)


# standardize values
df_classify$z.abv = scale(df_classify$beer.abv)
df_classify$z.ibu = scale(df_classify$beer.ibu)


train.indices = sample(1:dim(df_classify)[1],round(split.perc * dim(df_classify)[1]))

train = df_classify[train.indices,]
test = df_classify[-train.indices,]

dim(df_classify) # 939
dim(train) # 657
dim(test) #282



# original scale
classifications = knn(train[,c(4,5)],test[,c(4,5)], train$beer.ipa.vs.ale, prob = TRUE, k = 6)
# confusionMatrix(table(test$beer.ipa.vs.ale,classifications))

# standardized
classifications = knn(train[,c(16,17)],test[,c(16,17)], train$beer.ipa.vs.ale, prob = TRUE, k = 16)
confusionMatrix(table(test$beer.ipa.vs.ale,classifications))


```


## Loop for many k and one training / test partition
```{r}
acc = data.frame(accuracy = numeric(30), k = numeric(30))

for(i in 1:30)
{
  classifications = knn(train[,c(16,17)],test[,c(16,17)], train$beer.ipa.vs.ale, prob = TRUE, k = i)
  table(test$beer.ipa.vs.ale, classifications)
  CM = confusionMatrix(table(test$beer.ipa.vs.ale,classifications))
  acc$accuracy[i] = CM$overall[1]
  acc$k[i] = i
}

plot(acc$k,acc$accuracy, type = "l", xlab = "k", ylab="accuracy")

```


## Loop for many k and many training / test partitions
```{r}

set.seed(1234)
iterations = 100
numks = 30


masterAcc = matrix(nrow = iterations, ncol = numks)

for(j in 1:iterations)
{
  train.indices = sample(1:dim(df_classify)[1],round(split.perc * dim(df_classify)[1]))
  train = df_classify[train.indices,]
  test = df_classify[-train.indices,]
  for(i in 1:numks)
  {
    classifications = knn(train[,c(16,17)],test[,c(16,17)], train$beer.ipa.vs.ale, prob = TRUE, k = i)
    table(test$beer.ipa.vs.ale, classifications)
    CM = confusionMatrix(table(test$beer.ipa.vs.ale,classifications))
    masterAcc[j,i] = CM$overall[1]
  }
  
}

MeanAcc = colMeans(masterAcc)

plot(seq(1,numks,1),MeanAcc, type = "l")

which.max(MeanAcc)
max(MeanAcc)



```

